Analytics Without Parameter Tuning Considered Harmful?

### 2015-03-12 learners.tex
Updates from ShareLaTeX

Dta miners find generalizations (a.k.a. summaries)
of the data. All such generalizations must be smaller than the data it generalizes (otherwise we might as well just keep the original data and ignore the generalization)\footnote{Technical note:
even instance-based reasoners uses generalizations when they partially match new cases to old;
or if they use some  clustering algorithm to restrict the case matching
to just the subset of old cases most similar  to new  cases.}. 
Generalization is important since,
without generalization, all we can do is perform exact matches
of new situations to a database of old examples. This is not recommended practice since,
 if the new situation has not occurred  before, then nothing in the data  will
 matches the new situation
(and we cannot make any predictions).

The key point here is that there are many ways to summarize data and exceConsider some learner that
builds predictive rules form a subset of  attributes $A$ ranges. Assuming numeric attributes
divided into five ranges, then the 20 attributes of \fig{ck} can be combined into 
$2^{5*20}> 10^{30}$
possible rules

>>>>

Dta miners find generalizations (a.k.a. summaries)
of the data. All such generalizations must be smaller than the data it generalizes (otherwise we might as well just keep the original data and ignore the generalization)\footnote{Technical note:
even instance-based reasoners uses generalizations when they partially match new cases to old;
or if they use some  clustering algorithm to restrict the case matching
to just the subset of old cases most similar  to new  cases.}. 
Generalization is important since,
without generalization, all we can do is perform exact matches
of new situations against  a database of old examples. This is not recommended practice since,
 if the new situation has not occurred  before, then nothing in the data  will
 matches the new situation
(and we cannot make any predictions).

The key point here is that, at least is software engineering,
the raw data contains noise and superflous details (e.g. highly correleated attributes
or repeated/very similar rows). Recent results show

are many ways to summarize data.  A repeated result is
that a table of data with R rows and C columns can be reduced to around Consider some learner that
builds predictive rules form a subset of  attributes $A$ ranges. Assuming numeric attributes
divided into five ranges, then the 20 attributes of \fig{ck} can be combined into 
$2^{5*20}> 10^{30}$
possible rules

### 2015-03-12 tuningpredictor.tex
Updates from ShareLaTeX

\maketitle
\begin{abstract}
When learning
software defect predictors,   a
very simple optimizer called differential evolution needs can quickly
find  good tunings for that learner.
Such tunings can have a dramatic change to the performance of a learner;
e.g. in this paper we show that tuning can alter the precision of
a software defect predictor from 2\% to 98\%.

@@

One of the ``black arts'' of data mining is setting the tuning
parameters that control  the choices within that data miner.
Such tunings can have a dramatic change to the performance of a learner;
e.g. in this paper we show that tuning can alter the precision of
a software defect predictor from 2\% to 98\%.

Many researchers have applied automatic optimizers to find good tunings~\cite{cora10}  but not in the arean
of adjusting the parameters of data miners learning
static code defect predictors.
(to the best of our knowledge, this is the first such study).
This paper shows that for at least one software analytics problem
(software defect prediction from static code attributes),   a
very simple optimizer called 
{\em differential evolution} can  find good tunings.

The good news here is that that tuning defect predictors is very easy. Differential
evolution is so simple to implement that many
researchers just ``roll their own'' rather
than relying on some standard package.
By various measures
(presented below),  
tuning defect predictors is at least an order of magnitude
simpler than other standard optimization tasks. 
Further, measured in terms on runtimes, 
differental evolution takes just seconds
to a few minutes to propose highy effective tunings.

The not-so-good-news is that these results   challenge much prior work into defect prediction.
Several   prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task~.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
softwre defect predition~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests).

Another class of conclusions challenged by this paper are research papers
that use data miners to   show that certain factors
are more important than others for (say)
predicting defects. We show in this paper
that cuch conclusions
do not survive tuning. It turns out the after tuning,
different data sets show  that very different 
factors are ``best''. Also, even within the same data set, if we adjust the goals of the tuning then we arrive at different ``best'' factors.

Given this locality effect in what is ``best'',
then there are many   papers that may
need to be rewritten.
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted.

Similarly,   we should now revisit
all other conclusions that used data miners to rank different factors including:
\bi
\item
Bell et al.~\cite{bell2013limited} who say that
software defect prediction is not improved by adding attributes relating to the social
structure of programmer teams;
\item
Rahman et al
\cite{rahman2013how}  who say
 code metrics are  less useful than process
metrics for  predicting defects from static code analysis; 
\item Mosser et al.~\cite{moser2008comparative} who also say  that process metrics are better. 
\ei
There are many other examples of researchers
using data mining to claim that certain
factors are more important than others; e.g. ~\cite{zimmermann2007predicting,herzig2013predicting}.
All those results must  now be called into question.
At least for softwre defect prediction,
since is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.

The implications for other kinds of data mining tasks is unclear.
While this paper reports success with tuning defect predictors,
there are many of software analytics tasks 
(software development effort estimation, social network mining,
detecting duplicate issue reports, etc).
Since those other tasks often use the same kinds of learners
explored in this paper, it is clearly possible that
the conclusions of this paper apply to other tasks as well. 
Hence, we assert that it is now an open and pressing research question if
``analytics without parameter tuning is considered harmful?''.
Note that if the answer is ``yes'' then:
conferences like FSE
should routinely reject software analytics papers that do not
have a tuning study.

@@

One caveat before beginning- this paper explores
optimization via differential evolution.
XXX

>>>>

\maketitle
\begin{abstract}
One of the ``black arts'' of data mining is setting the tuning
parameters that control   choices within a data miner.  We offer a simple,
automatic, and very effective  method for finding those tunings.

For the purposes of learning
software defect predictors this optimization strategy can quickly
find  good tunings that  dramatically change   the performance of a learner.
For example,
in this paper we show   tunings that  alter deection  precision  
 from 2\% to 98\%.

@@

One of the ``black arts'' of data mining is setting the tuning
parameters that control  the choices within that data miner.
Many researchers have applied automatic optimizers to find good tunings.
For example, 
the search-based SE community has   developed techniques
for tuning that treating parameters  as a configuration
search space~\cite{cora10,Krogmann10}. More generally,
the field of {\em hyper-heuristics} explores methods for auto-adapting
options within the device searching for solutions~\cite{jia2013learning}.

To the best of our knowledge, this paper is the first exploration in software engineering
of applying automatic optimizers to the task of tuning data miners for defect prediction
(though see~\cite{cora10,balogh12} for work on effort estiamtion and genetic optimizers).
We shows that for at least one software analytics problem
(software defect prediction from static code attributes),   a
very simple optimizer called 
{\em differential evolution} can  find good tunings.

The good news here is that  tuning defect predictors is very easy and very effective. 
Differential
evolution is so simple to implement that many
researchers just ``roll their own'' rather
than relying on some standard package.
By various measures
(presented below),  
tuning defect predictors is at least an order of magnitude
simpler than other standard optimization tasks. 
Also, measured in terms on runtimes, 
differental evolution takes  seconds
to a just few minutes to propose highy effective tunings.
Further, the tunings found by differential evolution
can have a dramatic change to the performance of a learner;
e.g. in this paper we show that tuning can alter the precision of
a software defect predictor from 2\% to 98\%.

The not-so-good-news is that these results  challenge much prior work into defect prediction.
Several   prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task~.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
softwre defect predition~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests). Hence, all those prior results that ranked learners for software
analytics now need to be revisited and perhaps revised.

Another class of conclusions challenged by this paper are research papers
that use data miners to   show that certain factors
are more inflential than others for (say)
predicting defects. As shown below, such conclusions can be dramatically
changed by the tuning process since those  ``influential'' factors are very different pre- and post- tuning. Also, those factors tend to  change from project to project or if the goal
of the tuning is altered.
Hence, many old papers    need to be revisted  and perhaps revised~\cite{bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting}.  
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted since that
Menzies study did conduct a tuning study before reporting what factors the data miners
found where most influential.

In summary,it is now an open and pressing research issue to check if
analytics without parameter tuning is considered {\em harmful} or, at the 
very least, {\em misleading}.
At least for defect prediction,
it is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.
There are many other kinds of software analytics tasks 
(software development effort estimation, social network mining,
detecting duplicate issue reports, etc) and the implication of this
study for those tasks is unclear. 
However,  those other tasks often use the same kinds of learners
explored in this paper so it is clearly possible that
the conclusions of this paper apply to other SE analytics tasks as well.

@@

Two caveats before beginning. Firstly, we are {\em not} saying that {\em all learning requires
tuning}. For example, when proposing fixes to software, Weimar et al. use a genetic learner
to propose patches~\cite{Weimer:2009}. Suppose  the performance criteria for that work is how many known bugs can be found
and fixed automatically. Note that there is no {\em better than} in that performance criteria so
that study does not need tuning. However, once {\em better than} enters the performance criteria
then this becomes a race between competing methods or attributes.  In such a race, it is unfair
to hobble one competitor with poor tunings. Hence, we recommend tuning for  all the data mining
studies mentioned above\cite{lessmann2008benchmarking,hall11,me07b,bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting} since
they are arguing that one learner or one set of attributes
is {\em better than} another.

Secondly,
One caveat before beginning- this paper explores
optimization via differential evolution.
XXX
